Interpretable Neural Networks using EAGGA

### 2025-03-08 paper/main.tex
add software references
rename some headlines, add known software bugs (+ their circumvention)

\ShortHeadings{Extending EAGGA to Neural Networks}{Simon Stürzebecher}
\firstpageno{1}

@@

\title{Extending EAGGA to Neural Networks}

@@

% feels better in front of AutoML, as it feels more qualitative, less quantitative
\subsection{Explainable Artificial Intelligence and Interpretability}
\subsubsection{Why is interpretability desirable?}
\subsubsection{Practical implementations}
(1) post-hoc methods
  - give examples
  - why are they not perfectly suited
(2) model-based

\subsection{Automated Machine Learning (AutoML)}
\subsubsection{Hyperparameter Optimization (HPO)}
- grid + random search
- Bayesian optimisation
-> both unsuitable for our task
  + would be ok for weight decay (L2) + dropout optim
  + but: approaches for feature selection (L1) not feasible
    (a) 0-1 penalty on included features added to loss, problems:
      - how to choose regularisation parameter?
      - not all features equally important -> weighted 0-1 loss? how to choose weights?
    (b) directly optimise over binary hpyercube -> how to optim over non-ordinal discrete space?
-> evolutionary algorithms can be a solution for these types of problems
  * CMA-ES
  * NSGA-II
    - non-dominated sorting
    - crowding distance
    -> give intuition why non-dom-sort (asc) + crowding dist (desc) makes sense for selecting individuals
\subsubsection{Neural Architecture Search (NAS)}
not sure if super relevant as I don't use a specific NAS technology (e.g. one-shot, etc.) but simply optimise over \# layers + \# nodes as HPs

\subsection{Multi-Objective Optimization}
- problem: all aforementioned approaches are a-priori approaches -> we choose best HPs (weight decay, features used, dropout) w.r.t. performance
- intro MOO

@@

      * always train for min 200 epochs, keep track of model with lowest loss

@@

      * if no early stopping, train each model for max 10 minutes

@@

@@

\appendix
\section{}
\label{app:theorem}

>>>>

\ShortHeadings{Interpretable Neural Networks using EAGGA}{Simon Stürzebecher}
\firstpageno{1}

@@

\title{Interpretable Neural Networks using EAGGA}

@@

% feels better in front of AutoML, as it feels more qualitative, less quantitative
\subsection{Interpretability}
- Why Interpretability is desirable?
- Practical implementations
  (1) post-hoc methods
    - give examples
    - why are they not perfectly suited
  (2) model-based

\subsection{Hyperparameter Optimization (HPO)}
- grid + random search
- Bayesian optimisation
-> both unsuitable for our task
  + would be ok for weight decay (L2) + dropout optim
  + but: approaches for feature selection (L1) not feasible
    (a) 0-1 penalty on included features added to loss, problems:
      - how to choose regularisation parameter?
      - not all features equally important -> weighted 0-1 loss? how to choose weights?
    (b) directly optimise over binary hpyercube -> how to optim over non-ordinal discrete space?
-> evolutionary algorithms can be a solution for these types of problems
  * CMA-ES
  * NSGA-II
    - non-dominated sorting
    - crowding distance
    -> give intuition why non-dom-sort (asc) + crowding dist (desc) makes sense for selecting individuals
\subsection{Neural Architecture Search (NAS)}
not sure if super relevant as I don't use a specific NAS technology (e.g. one-shot, etc.) but simply optimise over \# layers + \# nodes as HPs

\subsection{Multi-Objective Optimization}
- problem: all aforementioned approaches are a-priori approaches -> we choose best HPs (weight decay, features used, dropout) w.r.t. performance
- intro MOO
\subsubsection{Scalarisation}
\subsubsection{Pareto-optimality}

@@

      * for each fold, always train for min 200 epochs, keep track of model with lowest loss

@@

      * if no early stopping, train each fold for max 10 minutes

@@

  * known bugs
    - in rare cases (anecdotally every ~2 datasets), gga\_mutate seems to be generating -1 as monotonicity attribute, despite np.random.randint(low=0, high=2, size=1)
      * loop crashes at group\_structure creation, for these cases subtract previous runtime from 8hrs + load from last generation in output via load\_population
      * so far only happened for madeline in after gen-8.json after 5hrs 45mins (at start of gen-9, which wasn't exported yet) -> loaded this and ran for another 2hrs 15mins

@@

\appendix
\section{Software used}
for implementation we used openml \cite{OpenML}, \cite{OpenMLPython}, numpy \cite{numpy}, pandas \cite{pandas1}, \cite{pandas2}, pytorch \textbf{TODO},
scikit-learn \cite{scikit-learn}, scipy \cite{SciPy}, pymoo \cite{pymoo}, and tqdm \cite{tqdm}

\section{}
\label{app:theorem}
